---
title: "Housing in Ames"
author: "Yanzun Jiang, Siyuan Lu, Yi Tang"
date: today
date-format: long
thanks: "Code and data supporting this proposal is available at: <https://github.com/Stary54264/Housing-in-Ames>"
format: pdf
number-sections: true
toc: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(AmesHousing)
library(ggplot2)
```

\newpage

# Introduction

@citeR

# Data Description

# Ethics Discussion

# Preliminary Results

```{r}
# Read data from cleaned data file
cleaned_data <- read_csv(here::here("data", "cleaned_data.csv"))

# Convert variables to factors

cleaned_data$central_air <- as.factor(cleaned_data$central_air)
cleaned_data$roof_style <- as.factor(cleaned_data$roof_style)

# Fit the linear model
model <- lm(sale_price ~ lot_area + overall_qual + year_built + roof_style + 
            mas_vnr_area + total_bsmt_sf + central_air + garage_area + misc_val, data = cleaned_data)

# Residual Plot
ggplot(cleaned_data, aes(x = fitted(model), y = resid(model))) +
  geom_point() +
  geom_hline(yintercept = 0, col = "red") +
  xlab("Fitted Values") +
  ylab("Residuals") +
  ggtitle("Residual Plot")

# QQ plot
qqnorm(resid(model))
qqline(resid(model), col = "red")

```

\newpage

\appendix

# Appendix

## Contributions

Yanzun Jiang: Organized discussions and meetings; set up Github workspace for collaborating; downloaded data for setting up the linear regression model from package `AmesHousing` (@ameshousing); cleaned data to make further analysis easier.

Siyuan Lu: 

Yi Tang: Build a linear regression model to predict house sale prices by using five key predictors in cleaned data. It assisted to understand the relationship between variables and ensure data meets key assumptions for statistical validity.

\newpage

# References
